---
title: "Yeting_challengeProject"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Coding Challenge Project

This challenge project is based on the following publication.

Multiomics modeling of the immunome, transcriptome, microbiome, proteome and metabolome adaptations during human pregnancy” by Ghaemi et al, Bioinformatics 2018.

Datasets was downloaded from https://nalab.stanford.edu/wp-content/uploads/pregnancymultiomics.zip.


```{r GetFamiliarwithData, eval=TRUE, include=TRUE}
load(file = "/Users/yetingzhang/Desktop/Paper-Code-01May2018/Data/Data.Rda")

#Take a look at the data and print them into text files for easy interpretation (small memory footpring on my personal laptop)
#write.table(featurepatients, file = "/Users/yetingzhang/Desktop/featurepatients.txt", quote = FALSE)
#write.table(featuretimes, file = "/Users/yetingzhang/Desktop/featuretime.txt", quote = FALSE)
#write.table(featureweeks, file = "/Users/yetingzhang/Desktop/featureweeks.txt", quote = FALSE)
#Take a look at RNAseq data
#write.table(InputData[[1]], file = "/Users/yetingzhang/Desktop/Inputdata_RNA.txt", quote = FALSE)
# Take a look at microbiome data
#write.table(InputData[[4]], file = "/Users/yetingzhang/Desktop/Inputdata_micbiome.txt", quote = FALSE)

```

## Questions to ask: Are there any outliers within the dataset?


```{r do PCA on some datasets, include=TRUE}
#look into microbiome data first, and see if there's any outliers identified, extract microbiome data
#install.packages("ggfortify")
library(ggfortify)
#install.packages("factoextra")
library(factoextra)
micro <- InputData[[4]]

# Look into postpartum data (serve as a control dataset)

fourth <- micro[52:68,]
# perform basic pca analysis
pca_fourth <- prcomp(fourth)
groups4th <- factor(row.names(fourth))

fviz_eig(pca_fourth)
# the scree plot shows that the top 2 pcas already explained the majority of the variances in the data

# plot pc1 vs pc2 using postpartum patients microbiome dataset
fviz_pca_ind(pca_fourth,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )

# Based on the postpartum data, PTLG016 looks more like an outlier.

# 1st trimester dataset (first 17 rows of the table)
first <- micro[1:17,]
# perform basic pca analysis using basic R package
pca_first <- prcomp(first)
groups <- factor(row.names(first))

#autoplot(prcomp(first), label = TRUE, label.size = 3)

fviz_eig(pca_first)
# the scree plot shows that the top 2 pcas already explained the majority of the variances in the data
fviz_pca_ind(pca_first,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )



# Based on the figure, PTLG016 and PTLG020 seems like outliers. Look into 2nd trimester

second <- micro[18:34,]
# perform basic pca analysis
pca_second <- prcomp(second)
groups2nd <- factor(row.names(second))

#autoplot(prcomp(second), label = TRUE, label.size = 3)
fviz_eig(pca_second)
# the scree plot shows that the top 2 pcas already explained the majority of the variances in the data
# plot pc1 vs pc2 using second trimester patients microbiome dataset
fviz_pca_ind(pca_second,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )


#PTLG020 cluster again stand out as an outlier based on second trimester data

#look into third trimester

third <- micro[35:51,]
# perform basic pca analysis
pca_third <- prcomp(third)
groups3rd <- factor(row.names(third))

fviz_eig(pca_third)
# the scree plot shows that the top 2 pcas already explained the majority of the variances in the data

# plot pc1 vs pc2 using third trimester patients microbiome dataset

fviz_pca_ind(pca_third,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )

#PTLG016 cluster again stand out as an outlier based on third trimester data


```


### Overall, PTLG016 is more like an outlier based on PCA results on 3 gestational stages/out of total 4 stages.